FGFR1/3 Signaling as Achilles’ Heel of Phenotypic Diversity in Urothelial Carcinoma

### 2025-03-20 report/markdown/rebuttal_letter.Rmd

---
title: "FGFR1/3 Signaling as Achilles’ Heel of Phenotypic Diversity in Urothelial Carcinoma"
subtitle: "Rebuttal Letter"
author: "Department of Urology, Medical University of Innsbruck"
date: "`r format(Sys.time(), '%Y-%m-%d')`"

output: 
  bookdown::html_document2:
    css: "style.css"

bibliography: blca.bib

csl: frontiers_medical.csl

header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}

---

```{r, setup, include = FALSE}

library(Cairo)

opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      dev = "cairo_pdf", 
                      dpi = 600)

set_flextable_defaults(font.family = "Cambria", 
                       font.size = 10)

```

\newpage

# Reviewer 1

## Point 1

### Issue

In some of the cohorts, there are patients with NMIBC (i.e. prior MSK IMPACT study with 20 - 25% NMIBC) and upper tract disease (i.e. IMvigor). 
There are prior studies suggesting genomic landscapes differ widely among NMIBC vs. MIBC and upper tract vs. bladder cancers. 
Therefore, it seems it would be beneficial to look at each of these groups separately (MIBC, NMIBC, and upper tract) rather than lump them all together. 
Can the authors provide more justification about why they have included NMIBC and upper tract disease in this manuscript instead of focusing just on MIBC?

### Response

Thanks for this constructive comment. 
We agree that urothelial tumors with different anatomical locations but also NMIBC and MIBC differ in their transcriptomic landscapes. 
Because of the brief report format of our manuscript, we decided against separate analyses of upper/lower tract malignancies and comparison of NMIBC and MIBC, and will consider them in our future works. 
Instead, we constrained the analyzed data sets to patient with MIBC of the bladder. 
If no NMIBC/MIBC information was provided, we restricted the cancer samples to T2 - T4 stage tumors. 
Please note that because of the very rudimentary clinical information for the GENIE cohort, we were not able to filter the samples by anatomical location and invasiveness. 
Yet, as evident from __Figure 1B__, there were no substantial differences in location and density of mutations of _FGFR3_ between the unfiltered GENIE cohort and the MSK IMPACT, TCGA BLCA and BCAN cohorts restricted to bladder-only MIBC.

The final numbers of investigated cancer samples were: 
n = `r length(genie$analysis_ids)` in the GENIE cohort, 
n = `r length(msk$analysis_ids)` in the MCK IMPACT cohort, 
n = `r length(tcga$analysis_ids)` in the TCGA BLCA cohort, 
n = `r length(imvigor$analysis_ids)` in the IMvigor, and 
n = `r length(bcan$analysis_ids)` in the BCAN collective.

## Point 2

### Issue

Were any of the samples pre-treated in the five cohorts used? 
What about for the drug screens, were any of those samples pre-treated?

### Response

This is certainly an important point.

Samples in the bulk cancer transcriptome cohorts were indeed obtained from patients with diverse treatment history. 
In __Supplementary Table S1__ of the revised manuscript, we present frequencies of patients with non-invasive intra-vesical treatment, surgery and distribution of surgery types, neoadjuvant treatment, adjuvant systemic treatment, systemic chemo- and immunotherapy, and radiation. 
For most of the bulk cancer data sets (GENIE, MSK IMPACT, IMvigor, BCAN), this information was published along with the molecular data or extracted from the original publications. 
For the TCGA BLCA cohort, the treatment information was derived from papers by Robertson et al. [@Robertson2017] and Moiso [@Moiso2021].

Concerning the cell lines, we extracted demographic and clinical information of the donors from the DepMap portal [@DepMap2024] and present it in __Reviewer Table \@ref(tab:tab-cell-lines)__. 
Unfortunately for none of the analyzed cell lines, treatment information was available.

## Point 3

### Issue

While the authors show they can accurately predict the consensus classifier with FGF/FGFR pathway genes - the next step is what to do with the consensus classifier information? Currently we do not make any therapeutic recommendations based on the groups, so its a bit unclear how helpful this is moving forward.

### Response

We are completely aware that, at the moment, there are no recommendations on diagnostic and treatment of the MIBC consensus classes. 
Following your suggestion, we predicted metrics of drug resistance (IC50 and AUC) for the bulk cancers in the TCGA BLCA, IMvigor, and BCAN cohorts with RIDGE linear models trained with drug response and transcriptome data of epithelial pan-cancer cell lines in the GDSC, CTRP2, and PRISM in vitro drug screening experiments [@Yang2013; @Seashore-Ludlow2015; @Corsello2020]. 
Of note, this machine learning strategy employed our development package [_htGLMNET_](https://github.com/PiotrTymoszuk/htGLMNET) and regularized linear modeling approach to drug resistance prediction proposed by Geeleher at al. [@Geeleher2014] and successfully employed in our recent publication on testicular cancer [@Torzsok2025]. 
Subsequently, we compared the predicted IC50 and AUC resistance metrics between the MIBC consensus classes (__Supplementary Figure XXX___). 
These new data shows that for most of pan-FGFR inhibitors (AZD4547, FGFR3831, FGFR0939, erdafitinib), good sensitivity was predicted for the LumP class, the predicted response was highly heterogeneous in the renaming subsets. 
In^particular, for Ba/Sq cancers resistance was forecasted, while stroma-rich were put forward as erdafitinib-responsive malignancies. 
We also compared the resistance metrics of FGFR-targeting drugs for cancers with and without mutation of _FGFR3_ (__Supplementary Figure XXX__); in this case, no differences in predicted response between WT and mutant specimens were observed.

## Point 4

### Issue

Does Kamoun et al. provide SHAP values for the 857 genes in their classifier? 
If so, would be interesting to know the value for FGF/FGFR genes. 
What about any of the other consensus classifiers?

### Response

We appreciate this point a lot. 
Unfortunately, were were not able to find SHAP variable importance measures for the nearest centroid classification tool by Kamoun and colleagues. 
As a workaround, we trained Elastic Net and Random Forest machine learning models of MIBC consensus classes in a training subset of two-thirds of observations of the TCGA BLCA cohort with the following explanatory factors:

* ComBat-adjusted $log_2$ expression levels of `r length(ml_globals$variables)` FGFR/FGF/FGFBP-related genes as outlined in the initial version of the manuscript

* Combat-adjusted $log_2$ expression levels of `r length(mibc_globals$variables)` MIBC class-defining genes proposed by Kamoun et al [@Kamoun2020] and shared by the TCGA BLCA, IMvigor, and BCAN expression data sets

In evaluation of model performance, the MIBC consensus class assignment by the genuine nearest centroid tool by Kamoun et al. served as the ground truth. 
As presented in __Supplementary Figure YYY__ predictions of consensus classes by the models with the full set of predictors by Kamoun et al. were emulated the original classification tool with very high accuracy (test collectives, accuracy: `r mibc_comp$stats %>% filter(model == 'mibc_genes', !cohort %in% c('tcga', 'tcga_train'), algorithm != 'ensemble') %>% .$Accuracy %>% signif(2) %>% range %>% paste(collapse = ' to ')`, Cohen's $\kappa$: `r mibc_comp$stats %>% filter(model == 'mibc_genes', !cohort %in% c('tcga', 'tcga_train'), algorithm != 'ensemble') %>% .$Kappa %>% signif(2) %>% range %>% paste(collapse = ' to ')`). 
Their performance was generally better that accuracy of the models employing solely the FGFR/FGF/FGFBP-related genes (test collectives, accuracy: `r mibc_comp$stats %>% filter(model == 'fgfr_genes', !cohort %in% c('tcga', 'tcga_train'), algorithm != 'ensemble') %>% .$Accuracy %>% signif(2) %>% range %>% paste(collapse = ' to ')`, Cohen's $\kappa$: `r mibc_comp$stats %>% filter(model == 'fgfr_genes', !cohort %in% c('tcga', 'tcga_train'), algorithm != 'ensemble') %>% .$Kappa %>% signif(2) %>% range %>% paste(collapse = ' to ')`), especially when it comes to assignment of cancer samples to LumU class. 
Please note, that this later observation corroborates findings presented in the initial manuscript, that, except of _KL_ gene, there are no good FGFR/FGF/FGFBP-related markers of LumU cancers.

For the Elastic Net and Random Forest models of the consensus MIBC classes using the full gene set by Kamoun et al. SHAP importance measures of the explanatory variables were computed. 
As shown in __Supplementary Figure ZZZZ__, except of _CD44_ whose levels were highly predictive for Ba/Sq cancers, none of the FGFR/FGF/FGFBP-related genes present in the genuine consensus class signature (`r intersect(ml_globals$variables, mibc_globals$variables) %>% map_chr(~paste0('_', .x, '_')) %>% collapse_and`) was found to be highly influential for the MIBC consensus class forecasts.

## Point 5

### Issue

For the doses used for the cell culture experiments, do we know which concentration corresponds to a level that we would see in patients when using erdafitinib?

### Response

Thank you for this important comment. 
In in vitro experiments we used 0 to 10 µM erdafitinib, which corresponds to around 0 to 4.5 mg/L erdafitinib. 
The recommended dose for urothelial carcinoma patients with susceptible genetic alterations in the FGFR is 8 mg orally once daily [@FDA2024]. 
Treating 2D cell lines cannot be compared directly to administration in humans due to the complex differences in pharmacokinetics. 
Therefore, more research, in particular in pre-clinical animal models, is needed that can translate our findings to the clinical setting. 
However, the erdafitinib concentration range in our study is commonly used in in vitro drug screenings, and has also been used in multiple reports on erdafitinib in urothelial cancer [@Ohtsu2022; @Li2023; @Zhong2024; @Zhong2024].

## Point 6

### Issue

Isn't it plausible there is something co-expressed with the FGFR pathways that could be the true oncogenic driver? 
Or be the real cause of the different phenotypes seen? 
I'm not sure this study can prove that is not the case so this may need to be acknowledged as a limitation.

### Response

We agree with you. 
By means of statistical inference and modeling, we cannot prove causality, i.e. unequivocally demonstrate that FGFR signaling drives bladder cancer in general. 
For this reason, we re-formulated carefully the title, abstract, and conclusions of the manuscript and abstain from the 'driver' term. 
Instead, we underline diversity of activation mechanisms of FGFR signaling in urothelial cancers and which lets us advocate for a broader use of FGFR blockers, independently of _FGFR3_ mutations.

Nevertheless, inspired by you comment, we sought to carefully gauge activity of FGFR signaling by comparison of scores [@Hanzelmann2013] of `r length(sub_sig$variables)` FGFR-related gene signatures from MSig database between the consensus classes of MIBC. 
Please note, that most of those signatures included ligand and co-receptor genes, and may be hence regarded as a proxy of ligand- and co-receptor-dependent FGFR signaling. 
Interestingly, `r sub_sig$hm_plots$tcga$data %>% filter(variable_subset == 'Stroma-rich', !duplicated(variable)) %>% nrow` of those signatures were found to be significantly upregulated in stroma-rich cancers in at least two out of TCGA BLCA, IMvigor, and BCAN cohorts (__Supplementary Figure VVV__, __Supplementary Table VVV__).

## Point 7

### Issue

I understand this is a brief correspondence, but there is no mention of the limitations of this work in the manuscript.

### Response

To address this point, we have added the limitation section to the manuscript. 
As the limitations we list:

* incompleteness of the clinical and multi-omics data for the GENIE and MSK IMPACT cohorts (lacking transcriptome data), which precluded analyses of transcriptome, gene signatures, and drug response predictions

* observational character of our study based on bioinformatic analyses, which require validation in a real-world clinical setting

# Reviewer 2

## Point 8

### Issue

As a small suggestion, Authors should only briefly describe and further discuss the significant clinical impact of FGFR inhibitors in the current clinical practice.

### Response

Thanks for this constructive suggestion. 
We appended the introduction with a section on clinical relevance and important ongoing studies on FGFR inhibitors in urothelial carcinoma.
We also discuss that there is already compelling evidence that also patients without FGFR alterations may benefit from treatment with FGFR inhibitors.

# Reviewer 3

## Point 9

### Issue

Limitations: could the Authors add possible limitations of this study?

### Response

Following your and Reviewer 1 suggestion we included a listing of the following limitations in the final part of the text:

* incompleteness of the clinical and multi-omics data for the GENIE and MSK IMPACT cohorts (lacking transcriptome data), which precluded analyses of transcriptome, gene signatures, and drug response predictions

* observational character of our study based on bioinformatic analyses, which require validation in a real-world clinical setting

\newpage

# Tables for Reviewers

```{r tab-cell-lines, tab.cap = "Demographic, clinical, and pathological characteristic of donors of the DepMap urothelial cancer cell lines. Numeric variables are presented as medians with interquartile ranges and ranges. Qualitative variables are shown as percentages of the categories."}

flextable::flextable(rev_tabs$cell_lines) %>% 
  set_widths(c(4.5, 5.5)) %>% 
  theme_vanilla

```

\newpage

# References

### 2025-04-28 report/markdown/figure1_legend.Rmd
Figure/Table/Supplement reshaping
Changes in composition of figures and tables requested by a reviewer. The whole content of tables is now in Supplementary Material. Parts of Figure 1 were moved to Supplementary Material

```{r fig-figure-1a, fig.width = figur::convert(fig$figure_1a, to = 'in')$w, fig.height = figur::convert(fig$figure_1a, to = 'in')$h, fig.cap = 'Alterations of FGFR-, FGF-, and FGFBP-coding genes in MIBC consensus molecular classes of urothelial cancer.'}

@@

__Figure 1A. Alterations of FGFR-, FGF-, and FGFBP-coding genes in MIBC consensus molecular classes of urothelial cancer.__

_Cancer samples from the TCGA BLCA, IMvigor, and BCAN data sets were assigned to the MIBC consensus molecular classes based on ComBat-adjusted whole-genome mRNA levels by the nearest centroid algorithm (R package consensusMIBC). Differences in frequency of somatic mutations and copy number variants of FGFR-, FGF-, and FGFBP-coding genes between the LumP (luminal papillary), LumU (luminal unstable), stroma-rich, and basal/squamous-like (Ba/Sq) consensus molecular classes were assessed by weighted permutation test. P values were corrected for multiple testing with the false discovery rate method. Presence/absence of selected genetic alterations in the consensus molecular classes in the TCGA BLCA cohort was visualized in a oncoplot. Alteration names (mut: mutation, amp: amplification, del: deletion) with effect sizes and p values of differences between the consensus classes are indicated in the Y axis. Significant effects are highlighted with bold font. Numbers of cancer samples in the consensus classes are displayed in the plot caption._

@@

```{r fig-figure-1b, fig.width = figur::convert(fig$figure_1b, to = 'in')$w, fig.height = figur::convert(fig$figure_1b, to = 'in')$h, fig.cap = 'Mutational hot spots of the FGFR3 gene.'}

@@

__Figure 1B. Mutational hot spots of the FGFR3 gene.__

_Frequency of mutations affecting residues of the FGFR3 protein was expressed as percentage of all cancer samples in the cohort and displayed in dot plots. Each point represents a single protein residue, point color codes for assignment of the residue to the protein domain. Candidate mutation hot spots with mutation rates $\geq$ 1% are highlighted with enlarged symbols and labeled with percentages and protein sequence changes. Cohort names and numbers of analyzed cancer samples are indicated in the plot facets. The protein domain scheme is displayed above the plot._

@@

```{r fig-figure-1c, fig.width = figur::convert(fig$figure_1c, to = 'in')$w, fig.height = figur::convert(fig$figure_1c, to = 'in')$h, fig.cap = 'Expression of genes coding FGFR, FGF, and FGFBP at mRNA and protein level.'}

@@

__Figure 1C. Expression of genes coding FGFR, FGF, and FGFBP at mRNA and protein level.__

_Levels of FGFR-, FGF-, and FGFBP-coding mRNA were determined in cancer tissue samples in the TCGA BLCA, IMvigor, and BCAN by RNA sequencing and published as $log_2$ gene counts (TCGA BLCA and IMvigor) or Z-scores of $log_2$ gene counts (BCAN). mRNA expression of the genes of interest in DepMap urothelial cancer cell lines was measured by RNA sequencing and provided as $log_2$ gene counts. Median Z-scores of $log_2$ mRNA expression are shown in a heat map. Each tile represents median expression Z-score of a gene in one cohort. Numbers of cancer samples are indicated in the X axis. White tiles represent genes not detected in each cohort. Protein expression of `r expl_expr$plot_data$hpa %>% filter(complete.cases(.)) %>% .$variable %>% unique %>% length` FGFR-, FGF-, and FGFBP-coding genes in the Human Protein Atlas (HPA) repository was quantified by immunohistochemistry (IHC). The protein expression is shown as IHC scores defined as a product of intensity (1: negative, 4: strong) and quantity of positive cells (1: none to 4: >75%). Median IHC scores with interquartile ranges are visualized as boxes with whiskers spanning over 150% of the interquartile ranges. Single IHC specimens are visualized as points. Numbers of IHC samples are indicated in the Y axis._

@@

```{r fig-figure-1d, fig.width = figur::convert(fig$figure_1d, to = 'in')$w, fig.height = figur::convert(fig$figure_1d, to = 'in')$h, fig.cap = 'Co-expression networks of FGFR-, FGF-, and FGFBP-coding genes.'}

@@

__Figure 1D. Co-expression networks of FGFR-, FGF-, and FGFBP-coding genes.__

_Co-expression networks were constructed in the TCGA BLCA, IMvigor, and BCAN cohorts for the FGFR-, FGF-, and FGFBP-coding genes, whose expression levels correlated with Spearman’s $\rho \geq$ 0.3. Isolated vertices of the networks were removed. The networks were visualized using the Fruchterman-Reingold algorithm. Points represent node vertices/genes, point shape codes for classification of the protein product. Point color represent hub scores, which measure the overall correlation strength. Lines represent edges, i.e correlations between expression levels with $\rho \geq$ 0.3. Line widths correspond to $\rho$ values. Note the central position, connectivity, and high hub scores of the genes coding for FGFR1, its ligands FGF2, FGF7 and FGF10, and the binding proteins SDC2, DCN, PTX3 and HSPG2 preserved in all cohorts. This suggests that ligand-dependent signaling via FGFR1 may serve as a substantial signaling pathway in urothelial cancer._

@@

```{r fig-figure-1e, fig.width = figur::convert(fig$figure_1e, to = 'in')$w, fig.height = figur::convert(fig$figure_1e, to = 'in')$h, fig.cap = 'Differential expression of FGFR-, FGF-, and FGFR-coding genes in the consensus molecular classes of urothelial cancers.'}

@@

__Figure 1E. Differential expression of FGFR-, FGF-, and FGFR-coding genes in the consensus molecular classes of urothelial cancers.__

_Differences in log2-transformed mRNA levels of FGFR-, FGF-, and FGFBP-coding genes between the LumP, LumU, stroma-rich, and Ba/Sq consensus molecular classes (Figure 1A) were assessed by one-way ANOVA with $\eta^2$ effect size statistic. Differences in $log_2$-transformed gene expression between a class and LumP were investigated by two-tailed T test. P values were corrected for multiple testing with the false discovery rate method (FDR). Significantly differentially regulated genes were defined by p(ANOVA) < 0.05, $\eta^2$ 0.06, pFDR(T test) < 0.05, and at least 1.25-fold up- or downregulation in a class as compared with LumP cancers. Z-scores of $log_2$ expression levels in the TCGA BLCA cohort of significantly differentially regulated genes shared by at least two cohorts are displayed in a heat map. Each tile represents a single cancer sample. Significant effects are labeled with bold font in the Y axes. Numbers of cancer samples in the consensus classes are indicated in the plot caption._

@@

```{r fig-figure-1f-elnet, fig.width = figur::convert(fig$figure_1f_elnet, to = 'in')$w, fig.height = figur::convert(fig$figure_1f_elnet, to = 'in')$h, fig.cap = 'Importance of explanatory variables in the Elastic Net model of MIBC molecular consensus classes of urothelial cancers assessed by the SHAP algorithm.'}

fig$figure_1f_elnet$plot

@@

__Figure 1F. Importance of explanatory variables in the Elastic Net model of MIBC molecular consensus classes of urothelial cancers assessed by the SHAP algorithm.__

_Contributions of the explanatory variables, i.e. the FGFR-, FGF-, and FGFBP-coding genes, to prediction of LumP, LumU, stroma-rich, and Ba/Sq consensus molecular classes by an Elastic Net machine learning model were estimated with the SHAP algorithm (Shapley additive explanations). SHAP values for single highly influential genes (top 10 highest mean absolute SHAP) and observations are visualized in violin/swarm plots. Each point represents a single observation, point color codes for the minimum/maximum-scaled $log_2$-transformed expression of the gene. The corresponding mean absolute SHAP values as metrics of overall gene importance are presented in bar plots. Colors of the bars code for the sign of association of SHAP values and gene expression (red: positive, blue: negative)._

```{r fig-figure-1f-ranger, fig.width = figur::convert(fig$figure_1f_ranger, to = 'in')$w, fig.height = figur::convert(fig$figure_1f_ranger, to = 'in')$h, fig.cap = 'Importance of explanatory variables in the Random Forest model of MIBC molecular consensus classes of urothelial cancers assessed by the SHAP algorithm.'}

fig$figure_1f_ranger$plot

__Figure 1F. Importance of explanatory variables in the Random Forest model of MIBC molecular consensus classes of urothelial cancers assessed by the SHAP algorithm.__

_Contributions of the explanatory variables, i.e. the FGFR-, FGF-, and FGFBP-coding genes, to prediction of LumP, LumU, stroma-rich, and Ba/Sq consensus molecular classes by a Random Forest machine learning model were estimated with the SHAP algorithm (Shapley additive explanations). SHAP values for single highly influential genes (top 10 highest mean absolute SHAP) and observations are visualized in violin/swarm plots. Each point represents a single observation, point color codes for the minimum/maximum-scaled $log_2$-transformed expression of the gene. The corresponding mean absolute SHAP values as metrics of overall gene importance are presented in bar plots. Colors of the bars code for the sign of association of SHAP values and gene expression (red: positive, blue: negative)._

```{r fig-resistance, fig.width = figur::convert(fig$figure_1g, to = 'in')$w, fig.height = figur::convert(fig$figure_1g, to = 'in')$h, fig.cap = 'Resistance and sensitivity to pan-FGFR inhibitors of DepMap cancer urothelial cell lines.'}

fig$figure_1g$plot

__Figure 1G. Resistance and sensitivity to pan-FGFR inhibitors of DepMap cancer urothelial cell lines.__

_Information on resistance of urothelial cancer cell lines to pan-FGFR inhibitors in the GDSC1/2 (n = `r expl_res$sensitive$gdsc %>% select(-sample_id, -FGFR3_mutation) %>% map(~complete.cases(.x)) %>% map_dbl(sum) %>% range %>% paste(collapse = ' to ')` cell lines) and PRISM (n = `r expl_res$sensitive$prism %>% select(-sample_id, -FGFR3_mutation) %>% map(~complete.cases(.x)) %>% map_dbl(sum) %>% range %>% paste(collapse = ' to ')`) drug screening experiments was extracted from the DepMap repository. The resistance values were expressed as IC50 (concentration resulting in 50% growth inhibition, GDSC) or AUC (area under the dose-response curve, relative to DMSO control, PRISM). Significant sensitivity for the GDSC data was defined as IC50 below 10 µM, which was the maximum concentration of the FGFR inhibitors. Significant sensitivity for the PRISM data was defined as AUC < 1, which indicates stronger growth inhibition by a FGFR inhibitor than the DMSO control. IC50 and AUC for doxorubicin are presented as a positive cytotoxicity controls. Median IC50 and AUC values for the FGFR inhibitors with interquartile ranges are depicted as boxes with whiskers spanning over 150% of the interquartile ranges. Points represent single cell lines. Point symbols codes for the FGFR3 gene mutation status, point color represent sensitivity and resistance. The sensitivity cutoffs are represented by dashed lines. FGFR3 wild-type cell lines RT-112, RT4, SW 780, CAL-29, and J82 showed particularly high sensitivity to multiple FGFR blockers in the GDSC1/2 experiments. FGFR3 wild-type cell lines `r expl_res$box_plots$prism$data %>% filter(stri_detect(variable, fixed = 'erdafi'), resistance < 0.9) %>%  arrange(resistance) %>% .$name_lab %>% collapse_and` demonstrated a very good sensitivity to erdafitinib in the PRISM assay (AUC: `r expl_res$box_plots$prism$data %>% filter(stri_detect(variable, fixed = 'erdafi'), resistance < 0.9) %>%  arrange(resistance) %>% .$resistance %>% range %>% signif(2) %>% paste(collapse = ' to ')`). By contrast, `r expl_res$sensitive$prism %>% filter(.data[['PRISM_BRD-K84868168-001-01-0_erdafitinib']] == 'resistant') %>% .$sample_id %>% exchange(depmap$cell_lines, key = 'sample_id', value = 'cell_line_name') %>% collapse_and` cell lines were found to be resistant to erdafitinib with AUC values ranging from `r expl_res$box_plots$prism$data %>% filter(stri_detect(variable, fixed = 'erdafi'), status == 'resistant') %>%  arrange(resistance) %>% .$resistance %>% range %>% signif(2) %>% paste(collapse = ' to ')`._

```{r fig-figure-1h, fig.width = figur::convert(fig$figure_1h, to = 'in')$w, fig.height = figur::convert(fig$figure_1h, to = 'in')$h, fig.cap = 'Confluence of cultured FGFR1-4 wild-type (UM-UC-3, 5637 and RT112) and FGFR3 mutated (UM-UC-14 and UM-UC-6) UC cell lines after treatment with different concentrations of erdafitinib.'}

fig$figure_1h$plot

__Figure 1H. Confluence of cultured FGFR1-4 wild-type (UM-UC-3, 5637 and RT112) and FGFR3 mutated (UM-UC-14 and UM-UC-6) UC cell lines after treatment with different concentrations of erdafitinib.__

>>>>

```{r fig-figure-1a, fig.width = figur::convert(fig$figure_1a, to = 'in')$w, fig.height = figur::convert(fig$figure_1a, to = 'in')$h, fig.cap = 'Alterations of FGFR/FGF/FGFBP-coding genes in consensus molecular classes of MIBC.'}

@@

__Figure 1A. Alterations of FGFR/FGF/FGFBP-coding genes in consensus molecular classes of MIBC.__

_Differences in frequency of somatic mutations and copy number variants of FGFR/FGF/FGFBP-coding genes between the LumP (luminal papillary), LumU (luminal unstable), stroma-rich, and basal/squamous-like (Ba/Sq) consensus molecular classes of MIBC in the TCGA, BCAN and IMvigor cohorts were assessed by false discovery rate (FDR) corrected weighted permutation test. Presence/absence of selected genetic alterations (mut: mutation, amp: amplification, del: deletion) in the TCGA BLCA cohort was visualized in a oncoplot. Significant effects are highlighted with bold font. Numbers of cancer samples are displayed in the plot caption._

@@

```{r fig-figure-1b, fig.width = figur::convert(fig$figure_1b, to = 'in')$w, fig.height = figur::convert(fig$figure_1b, to = 'in')$h, fig.cap = 'Co-expression networks of FGFR/FGF/FGFBP-coding genes.'}

@@

__Figure 1B. Co-expression networks of FGFR/FGF/FGFBP-coding genes.__

_Co-expression networks were constructed for the FGFR/FGF/FGFBP-coding genes, whose expression levels correlated with Spearman’s $\rho \geq$ 0.3 and visualized for the TCGA BLCA cohort with the Fruchterman-Reingold algorithm. Points represent genes, point shape codes for the protein product type. Point color represent hub scores. Lines represent edges, i.e correlations with $\rho \geq$ 0.3. Line widths correspond to $\rho$ values._

@@

```{r fig-figure-1c, fig.width = figur::convert(fig$figure_1c, to = 'in')$w, fig.height = figur::convert(fig$figure_1c, to = 'in')$h, fig.cap = 'Differential expression of FGFR/FGF/FGFR-coding genes in consensus molecular classes of MIBC.'}

@@

__Figure 1C. Differential expression of FGFR/FGF/FGFR-coding genes in consensus molecular classes of MIBC.__

_Differences in $log_2$ mRNA levels of FGFR/FGF/FGFBP-coding genes between the LumP, LumU, stroma-rich and Ba/Sq consensus molecular classes were assessed by FDR-corrected one-way ANOVA with $\eta^2$ effect size statistic and two-tailed post-hoc T test. Differentially regulated genes were defined by p(ANOVA) < 0.05, $\eta^2 \geq$ 0.14, pFDR(T test) < 0.05, and at least 1.25-fold regulation in a class as compared with LumP cancers. Z-scores of $log_2$ expression levels of differentially regulated genes shared by at least two cohorts are displayed in a heat map for TCGA BLCA cancers. Significant effects are labeled with bold font. Numbers of samples are indicated in the plot caption._

@@

```{r fig-figure-1d, fig.width = figur::convert(fig$figure_1d, to = 'in')$w, fig.height = figur::convert(fig$figure_1d, to = 'in')$h, fig.cap = 'SHAP variable importance in the Elastic Net model of consensus molecular classes of MIBC with FGFR/FGF/FGFBP-coding genes as explanatory factors.'}

@@

__Figure 1D. SHAP variable importance in the Elastic Net model of consensus molecular classes of MIBC with FGFR/FGF/FGFBP-coding genes as explanatory factors.__

_Contributions of the FGFR-, FGF-, and FGFBP-coding genes to prediction of LumP, LumU, stroma-rich, and Ba/Sq consensus molecular classes of MIBC by the Elastic Net model were estimated with the SHAP algorithm (Shapley additive explanations). SHAP values for highly influential genes and observations are visualized in violin/swarm plots. Each point represents a single observation, point color codes for the minimum/maximum-scaled $log_2$-transformed gene expression. Mean absolute SHAP values as metrics of overall gene importance are presented in bar plots. Bar colors code for association of SHAP values and gene expression (red: positive, blue: negative)._

@@

```{r fig-resistance, fig.width = figur::convert(fig$figure_1e, to = 'in')$w, fig.height = figur::convert(fig$figure_1e, to = 'in')$h, fig.cap = 'Sensitivity to pan-FGFR inhibitors of DepMap cancer urothelial cell lines.'}

@@

__Figure 1E. Sensitivity to pan-FGFR inhibitors of DepMap urothelial cancer cell lines.__

_Resistance values in the PRISM (n = `r expl_res$sensitive$prism %>% select(-sample_id, -FGFR3_mutation) %>% map(~complete.cases(.x)) %>% map_dbl(sum) %>% range %>% paste(collapse = ' to ')`) drug screening experiment were expressed as AUC (area under the dose-response curve). Significant sensitivity for was defined as AUC < 1, which indicates stronger growth inhibition than the DMSO control. Doxorubicin is presented as a positive cytotoxicity control. Median AUC values with interquartile ranges are depicted as boxes with whiskers spanning over 150% of the interquartile ranges. Points represent single cell lines. Point symbols code for FGFR3 mutation, point color represents resistance/sensitivity. The sensitivity cutoff is represented by the dashed line._

@@

```{r fig-figure-1f, fig.width = figur::convert(fig$figure_1f, to = 'in')$w, fig.height = figur::convert(fig$figure_1f, to = 'in')$h, fig.cap = 'Confluence of cultured FGFR1-4 wild-type (UM-UC-3, 5637 and RT112) and FGFR3 mutated (UM-UC-14 and UM-UC-6) UC cell lines after treatment with different concentrations of erdafitinib.'}

fig$figure_1f$plot

@@

__Figure \@ref(fig:fig-figure-1f). Confluence of cultured FGFR1-4 wild-type (UM-UC-3, 5637 and RT112) and FGFR3 mutated (UM-UC-14 and UM-UC-6) UC cell lines after treatment with different concentrations of erdafitinib.__
